Model

### 2019-04-20 stats/model.Rmd

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

@@

```{r rsa model}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior); 
    var lexicon = lexiconObject[lexiconName];
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    if ("label" in utterance) {
      var truthValue = lexicon(utterance, obj, sem_knowledge);
      condition(truthValue)
    }
    return obj.shape 
  }})
})

var speaker = cache(function(obj, lexiconName, priorProbs, speakerOptimality, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var utterance = utterancePrior();
    var L0 = literalListener(utterance, priorProbs, sem_knowledge);
    factor(speakerOptimality * L0.score(obj.shape))
    return utterance
  }})
})

var pragmaticListener = cache(function(utterance, priorProbs, speakerOptimality, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior);
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    var S1 = speaker(obj, lexiconName, priorProbs, speakerOptimality, sem_knowledge);
    observe(S1, utterance)
    return obj.shape == "novel_object" ? 1 : 0
  }})
})

@@

  var params = map(function(cndtn){

@@

      var modelPredictions = pragmaticListener({label: "novel_word"}, priorProbs, speakerOptimality,sem_knowledge)

      observe(modelPredictions, row.correct)

@@

    return extend([
      {sem_int: sem_int, sem_slope: sem_slope, familiar: cndtn, so_int: so_int, so_slope: so_slope}
    ])

@@

params

}
'
```

@@

sem_know_model <- readRDS("../saves/sem_know_model.rds")

sem_know <- sem_know_model %>%
  select(value) %>%
  map_df(bind_rows)%>%
  rename(sem_int = `0.sem_int`,
         sem_slope = `0.sem_slope`,
         familiar = `0.familiar`,
         so_int = `0.so_int`,
         so_slope = `0.so_slope`)%>% 
  gather(parameter, value, -familiar)

sem_know_params <- sem_know %>%
  group_by(familiar,parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))
```

```{r}
ggplot(sem_know, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(parameter ~familiar)+
  theme_few()
```
## Plotting semantic knowledge
```{r}

@@

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% sem_know_params$familiar )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         familiar = Word)%>%
  select(familiar,mean_aoa)

plot_sem_know <- sem_know_params%>%
  select(-uci,-lci)%>%
  spread(parameter, -familiar)%>%
  expand(familiar, sem_slope, sem_int,age = me_model_data$age)%>%
  mutate(sem_know = plogis(sem_int + sem_slope * age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

ggplot(plot_sem_know, aes(x = age+2, y= sem_know, col = familiar))+
  geom_line(size = 1)+
  ylab("Semantic knowledge")+
  xlab("Age")+
  ylim(0,1)+
  theme_few()+
  scale_colour_viridis_d(name = "familiar object")
```

>>>>

```{r}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

@@

```{r rsa model}
rsaModel <- '
var literalListener = cache(function(utterance, priorProbs, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior); 
    var lexicon = lexiconObject[lexiconName];
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    if ("label" in utterance) {
      var truthValue = lexicon(utterance, obj, sem_knowledge);
      condition(truthValue)
    }
    return obj.shape 
  }})
}, 10000)

var speaker = cache(function(obj, lexiconName, priorProbs, speakerOptimality, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var utterance = utterancePrior();
    var L0 = literalListener(utterance, priorProbs, sem_knowledge);
    factor(speakerOptimality * L0.score(obj.shape))
    return utterance
  }})
}, 10000)

var pragmaticListener = cache(function(utterance, priorProbs, speakerOptimality, sem_knowledge){
  Infer({method: "enumerate", model: function(){
    var lexiconName = sample(LexiconPrior);
    var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
    var S1 = speaker(obj, lexiconName, priorProbs, speakerOptimality, sem_knowledge);
    observe(S1, utterance)
    return obj.shape == "novel_object" ? 1 : 0
  }})
}, 10000)

@@

  foreach(function(cndtn){

@@

      var modelPredictions = pragmaticListener({label: "novel_word"}, priorProbs, speakerOptimality,sem_knowledge)
      // display(modelPredictions.score(row.correct))
      observe(modelPredictions, row.correct)

@@

    query.add(cndtn, [sem_int, sem_slope])

@@

  query.add("speaker_optimality", [so_int, so_slope])
  return query 
}
'
```

@@

#sem_know_model <- readRDS("../saves/sem_know_model.rds")

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
  data = list(data = me_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH", 
                        samples = 5000, 
                        burn = 2500,
                        verbose = T,
                        verboseLag = 500)
)
```

```{r}

sem_know <- sem_know_model %>%
  filter(Parameter != "speaker_optimality") %>%
  mutate(Parameter = factor(Parameter)) %>%
  separate(value, into = c("intercept", "slope"), 
           sep = ", ") %>%
  mutate(intercept = as.numeric(str_remove_all(intercept, "[(c]")),
         slope = as.numeric(str_remove(slope, "[)]"))) %>%
  rename(item = Parameter) %>%
  gather(parameter, value, intercept, slope)

# sem_know <- sem_know_model %>%
#   select(value) %>%
#   map_df(bind_rows)%>%
#   rename(sem_int = `0.sem_int`,
#          sem_slope = `0.sem_slope`,
#          familiar = `0.familiar`,
#          so_int = `0.so_int`,
#          so_slope = `0.so_slope`)%>% 
#   gather(parameter, value, -familiar)

sem_know_params <- sem_know %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))
```

```{r}
ggplot(sem_know, aes(value))+
  geom_density()+
  xlab("Parameter")+
  facet_grid(parameter~item, scales = 'free')+
  theme_few()
```
## Plotting semantic knowledge
```{r}

@@

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% sem_know_params$item )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

plot_sem_know <- sem_know_params%>%
  select(-uci,-lci)%>%
  spread(parameter, -item) %>%
  left_join(expand.grid(
      age = unique(me_model_data$age),
      item = me_model_data$familiar), .) %>%
  mutate(sem_know = plogis(intercept + slope * age)) %>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

ggplot(plot_sem_know, aes(x = age+2, y= sem_know, col = item, group = item))+
  geom_line(size = 1)+
  ylab("Semantic knowledge")+
  xlab("Age")+
  ylim(0,1)+
  theme_few()+
  scale_colour_viridis_d(name = "familiar object")
```
